Search techniques for gaussian mixture modelling with Minimum Message Length, with applications to the origin of our Galaxy

### 2018-04-19 article/ms.tex
wip

Developing similar theoretical bounds for $I_2$ is not straightforward, because
the determinants of the covariance matrices of Gaussian components will not
-- to our knowledge -- strictly follow some expected distribution, in the way 
that the weights $\weights$ follow a multinomial distribution. Similarly, the
\emph{scale} of $|\veccov_k|$ will depend entirely on the distribution of the
data. In that sense, we cannot a priori state with any certainty about what 
values $|\veccov_k|$ might take (or $\log{|\veccov_k|}$) without knowing
something about the data first. Even demanding that the covariance matrices
$\veccov_k$ are positive semi-definite only informs us that $\log{|\veccov_k|}$
is finite. However, there are quantitative predictions we can make on the 
bounds of $\sum_{k=1}^{K}\log{|\veccov_k|}$ if something is known about the
data. Specifically if we consider the determinant of the sample covariance 
matrix, which we denote $|\cov_{s}|$, then we can state with confidence that
the sample covariance matrix describes the maximum amount of variance that 
could be captured by the components in the mixture. For example, a limiting
case of $\sum_{k=1}^{K}\log{|\veccov_k|}$ can be described when the determinant
of the sample covariance matrix must be equally (uniformly) captured by $K$ 
components in a mixture. In other words, the maximum variance in the data 
must be equally spread between the components in the mixture. Uniformly 
describing the variance in the data is the limiting case because any 
non-uniformity would lead to smaller $|\veccov|$ for one mixture (and larger 
for another), and that asymmetry would lead to a higher $I_2$ because 
$\log(x - \epsilon) + \log(x + \epsilon) < 2\log{x}$, and because of the 
$-(1 + \frac{D}{2})$ term that enters multiplicatively in the calculation of 
$I_2$. Thus, we can specify a lower bound on $I_2$ based on the determinant of
the sample covariance matrix.

>>>>

Developing similar theoretical bounds for $I_2$ is not straightforward, because
the determinants of the covariance matrices of Gaussian components will not
-- to our knowledge -- strictly follow some expected distribution, in the way 
that the weights $\weights$ follow a multinomial distribution. Similarly, the
\emph{scale} of $|\veccov_k|$ will depend entirely on the distribution of the
data. In that sense, we cannot a priori state with any certainty about what 
values $|\veccov_k|$ might take (or $\log{|\veccov_k|}$) without knowing
something about the data first. Even demanding that the covariance matrices
$\veccov_k$ are positive semi-definite only informs us that $\log{|\veccov_k|}$
is finite. However, there are quantitative predictions we can make on the 
bounds of $\sum_{k=1}^{K}\log{|\veccov_k|}$ if something is known about the
data. Specifically if we consider the determinant of the sample covariance 
matrix, which we denote $|\cov_{s}|$, then we can state with confidence that
the sample covariance matrix describes the maximum amount of variance that 
could be captured by the components in the mixture. For example, a limiting
case of $\sum_{k=1}^{K}\log{|\veccov_k|}$ can be described when the determinant
of the sample covariance matrix must be equally (uniformly) captured by $K$ 
components in a mixture. In other words, the maximum variance in the data 
must be equally spread between the components in the mixture. Uniformly 
describing the variance in the data is the limiting case because any 
non-uniformity would lead to smaller $|\veccov|$ for one mixture (and larger 
for another), and that asymmetry would lead to a higher $I_2$ because 
$\log(x - \epsilon) + \log(x + \epsilon) < 2\log{x}$, and because of the 
$-(1 + \frac{D}{2})$ term that enters multiplicatively in the calculation of 
$I_2$. Thus, we can specify a lower bound on $I_2$ based on the determinant of
the sample covariance matrix $|\cov_{s}|$.

The other extreme on $I_2$ can be approximated by estimating the \emph{minimum}
possible variance that could be captured by any single component.
